---
title: "Feb19-2020_SIO-RUsers_ggplot2-ggmap"
author: "Sabah Ul-Hasan"
output: pdf_document
---

How can we use R, R markdown, ggplot2, and ggmap to plot data on a publish-ready map? 
Code repo    https://github.com/almlab/analysis-salt-responsive/tree/master/analysis 
*Data from Malloy and Ul-Hasan et al (in progress)

Lines 15-52      Relevant package installation, and library load
Lines 54-154     Map demo: Figure 1 (a-b)
Lines 156-     Build on Figure 1 by R script (or own data) 

Install Relevant Packages and Load Libraries 
```{r}

##### Package installation #####
# You can uncomment these packages to re-install them and update the versions
# Note to check your version of R + RStudio when running into installation trouble

#  Need this package for downloading data directly from Github
## install.packages("RCurl") 
## packageVersion("RCurl") # v1.98.1.1 
### Helpful to know which version used for reproducibility (like lab notebook)

# Need packages for map coordinates
## if(!requireNamespace("devtools")) install.packages("devtools")
## devtools::install_github("dkahle/ggmap")
## packageVersion("ggmap") # v3.0.0
## install.packages("rgdal")
## packageVersion("rgdal") # v1.4.8

# Need this package for visualization of data 
## install.packages("ggplot2") 
## packageVersion("ggplot2") # v3.2.1

# Need this package to compile images as a unified figure
## install.packages("cowplot")
## packageVersion("cowplot") # v1.0.0
##### Package installation #####

# Can load directly, if packages already installed 
##### Load libraries #####
library("RCurl")
library("rgdal")
library("ggmap")
library("ggplot2")
library("cowplot") 
##### Load libraries #####

```

Create a map showing where samples were collected 
```{r}

##### Upload Data #####
# Load data from our Github repo as df
df <- read.csv(text=getURL("https://raw.githubusercontent.com/sabahzero/dataviz/master/Feb19-2020/SupplementalTable3_Data-Clean.csv")) 
summary(df) # Quick view of our data, we can also see this in our 'Environment'

# About the data:
## These are a collection of Californiconus californicus (California Cone Snail) shells from 1892 - 2018, and their associated measurements in prep for the manuscript with running title, "Bergmann’s Rule across the Pleistocene, Holocene, and Anthropocene in an Eastern Pacific neogastropod, Californiconus californicus (Reeve, 1844) (Gastropoda: Conidae)" 
### ME Malloy*, S Ul-Hasan*, LL Lewis, JKK Hofmeister, S Crickenberger, CA O'Leary, S Sindi, A Hendy, WF Gilly, T Woyke, CJ Nobile, LT Groves, JE Vendetti 
### https://github.com/MichaelMalloy/C_californicus_Morphology 
### Ch I of dissertation: https://escholarship.org/uc/item/7rn287kn 

# For a map, what about the data do we care about?
## Latitude, Longitude, and Epoch (Anthropocene, Holocene, or Pleistocene)
## shape files for ecoregions from http://www.marineregions.org/ (specifics in repo)
NC_poly=readOGR( "northern_ecoregions" , layer="ecoregions") 
SCB_poly=readOGR( "southern_ecoregions" , layer="ecoregions") 
MT_poly=readOGR( "magdaelena_ecoregions" , layer="ecoregions") 
##### Upload Data #####

##### Create Map (Figure 1a-b) #####
#  Need a personalized account with API key (map key) for Google maps
## Instructions: https://github.com/dkahle/ggmap
## ggmap::register_google(key = " ") 
### More helpful info: https://www.littlemissdata.com/blog/maps 

# Figure 1a: Larger map
## Parameters for All points
All_bbox <- make_bbox(lat = Latitude, lon = Longitude, data = df)
All_bbox # Informs us what to set for parameters
##     left        bottom    right       top 
##     -122.786    27.359    -114.734    37.721 
All=get_map(location=c(-129,20,-105,42), # adjust to include regions
            zoom = 5, 
            maptype = "terrain") # can adjust this type as well
## Create box for zoom of SoCal
SoCal_box = data.frame(x=c(-121,-121,-116.5,-116.5,-121), 
       y=c(35,32,32,35,35)) 
## Plot map 
Figure1a = ggmap(All) + # Google map of parameters
  geom_polygon(data = NC_poly, aes(x = long, y = lat), # polygons (shape files)
               fill = "white", alpha = 0.5) + 
  geom_polygon(data = SCB_poly, aes(x = long, y = lat),
               fill = "blue", alpha = 0.3) +
  geom_polygon(data = MT_poly, aes(x = long, y = lat),
               fill = "#666666", alpha = 0.5) + # grey
  stat_density_2d(data = df, # hypothetical population density
                  aes(x = Longitude,
                      y = Latitude,
                      fill = stat(level)),
                  alpha = .25,
                  bins = 250, # can adjust this for resolution
                  geom = "polygon") +
  scale_fill_gradientn(colors = "black") + # fills density layers as black 
  scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) + # colors by Epoch
  geom_point(data = df, aes(x = Longitude, y = Latitude, color = Epoch), 
             alpha = 0.15,
             size = 1) + # layers points
  geom_path(data=SoCal_box, aes(x,y), color="black", lwd=1) + # insert box
  labs(x="Longitude", y="Latitude") +
  theme(legend.position="none")
Figure1a

# Figure 1b: Zoom-in of SoCal
SoCal=get_map(location=c(-121,32,-116.5,35), # adjust based on SoCal_box
            maptype = "terrain-background") # can adjust this type as well
## Plot map
Figure1b = ggmap(SoCal) +
  geom_polygon(data = NC_poly, aes(x = long, y = lat),
               fill = "white", alpha = 0.5) + # white
  geom_polygon(data = SCB_poly, aes(x = long, y = lat),
               fill = "blue", alpha = 0.3) + # blue
    stat_density_2d(data = df, # density
                  aes(x = Longitude,
                      y = Latitude,
                      fill = stat(level)),
                  alpha = .25,
                  bins = 250,
                  geom = "polygon") +
  scale_fill_gradientn(colors = "black") + # color fill, black
  scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  geom_point(data = df, aes(x = Longitude, y = Latitude, color = Epoch), 
             alpha = 0.15,
             size = 1) +
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())
Figure1b # Note the ecoregions cut-out if polygon not fully included

# Grid using cowplot, and save
Figure1 = plot_grid(Figure1a, Figure1b, labels = c("A", "B"), label_size=12)
Figure1   

ggsave("Figure1.png", dpi = 800) 

##### Create Map (Figure 1a-b) #####

# What did we *not* do that's still left? 
## Group breakouts or work on own data

```

Exercises: Build on Figure 1 (or own data)
```{r}

# Now that you're in your group, copy/paste the script above and use this space to improve the script and/or image

## Select one of the following tasks (or try it out with your own data)
### 1. Remove the superfluous legend (or make it more informative if keeping)
### 2. Adjust the spacing of the legend to better balance out Figure 1
### 3. Adjust the map sizes such that there is less 'useless' space within A and B
### 4. Add a title to the figure
### 5. Make the script more DRY (ie a for loop to automate) 

## When you're done, save your .rmd script and .png image each as "Figure1_[First-LastName].[rmd or png]" 
## Then, upload onto the repo or email to sulhasan@scripps.edu to be shared
### You're also welcome to share insights from your own data, as you prefer!

```
